Open-source tools in R for landscape ecology

### 2021-01-27 paper/paper.Rmd
Final update survey data

```{r data-model, include = FALSE}
# data model
data_model <- dplyr::group_by(answers, data_model) %>%
  dplyr::summarise(n_rel = n() / n_people * 100) %>%
  dplyr::mutate(n_rel = round(n_rel, digits = digits)) %>%
  dplyr::arrange(-n_rel)

@@

@@

ggplot_packages <- ggplot(data = r_packages_df) +
  geom_bar(aes(y = r_package_reclass, x = n_rel, fill = col), stat = "identity") +
  # geom_text(aes(y = r_package_reclass, x = n_rel, label = label, col = col),
  #           nudge_x = 2, size = 2.5) +
  theme_classic(base_size = base_size) +
  theme(plot.title.position = "plot") +
  scale_x_continuous(breaks = seq(0, 35, 5), limits = c(0, 35)) +
  scale_fill_manual(values = c("grey50", "black")) +
  scale_colour_manual(values = c("grey50", "black")) +
  guides(fill = FALSE, colour = FALSE) +
  labs(x = "Number of package mentioned [%]", y = NULL,
       title = "List the three R packages related\nto landscape ecology you use the most")
```

@@

# no answer, no missing packages
id_na <- c("not sure", "not sure right now",
           "i have no additional needs for my current workflow", "none that i can think of",
           "i would like to know", "none", "i still have to fully understand what is there",
           "i don't know, sorry!", "r havent fail me yet", "na",
           "i'm not missing anything at present.", "não posso responder",
           "no", "unsure", "have not yet identified any.", "no package missing",
           "not enough overview to answer this", "?", "i am currently missing no methods",
           ";", "nothing yet", "i'm thinking", "nil")

>>>>

```{r data-model, include = FALSE}
# data model
data_model <- dplyr::group_by(answers, data_model) %>%
  dplyr::summarise(n_rel = n() / n_people * 100) %>%
  dplyr::mutate(n_rel = round(n_rel, digits = digits)) %>%
  dplyr::arrange(-n_rel)
```

@@

# remove escape characters
r_packages <- str_remove_all(r_packages, pattern = "\"")

@@

ggplot_packages <- ggplot(data = r_packages_df) +
  geom_bar(aes(y = r_package_reclass, x = n_rel, fill = col), stat = "identity") +
  # geom_text(aes(y = r_package_reclass, x = n_rel, label = label, col = col),
  #           nudge_x = 2, size = 2.5) +
  theme_classic(base_size = base_size) +
  theme(plot.title.position = "plot") +
  scale_x_continuous(breaks = seq(0, 35, 5), limits = c(0, 36)) +
  scale_fill_manual(values = c("grey50", "black")) +
  scale_colour_manual(values = c("grey50", "black")) +
  guides(fill = FALSE, colour = FALSE) +
  labs(x = "Number of package mentioned [%]", y = NULL,
       title = "List the three R packages related\nto landscape ecology you use the most")
```

@@

# no answer, no missing packages
id_na <- c("not sure", "not sure right now",
           "i have no additional needs for my current workflow", "none that i can think of",
           "i would like to know", "none", "i still have to fully understand what is there",
           "i don't know, sorry!", "r havent fail me yet", "na",
           "i'm not missing anything at present.", "não posso responder",
           "no", "unsure", "have not yet identified any.", "no package missing",
           "not enough overview to answer this", "?", "i am currently missing no methods",
           ";", "nothing yet", "i'm thinking", "nil", "none", 
           "nothing really but prefer to do programming in python", "i dont miss any")

### 2021-01-27 paper/paper.tex
Final update survey data

To better understand how the landscape ecology community uses \emph{R}, we conducted a short survey using mailing lists and social media to reach the community.
In total, the survey was answered by 103 participants, of which the majority were either ``PhD students'' (34\%), followed by ``Post-Docs'' (28.2\%) and ``Professors'' (12.6\%).
Other, less frequent answers were ``Data scientists,'' ``None of the above,'' ``Government employees'' ``Master's degree student'' and ``Bachelor's degree student'' (in decreasing order).

Most people use \emph{R} either ``daily'' (54.4\%) or a ``few times a week'' (36.9\%).
Almost half of all participants described themselves as ``advanced'' users (46.6\%), while 40.8\% described themselves as ``intermediate'' users.
Related to this, about half of the participants either implemented their own package (21.4\%) or plan to do so in the future (23.3\%) and most of these packages are hosted on \emph{GitHub} and/or \emph{CRAN}.

@@

More people use the raster data model (72.8\%) in comparison to the vector data model (27.2\%).
This was also represented in the most used \emph{R} packages (Fig. \ref{fig:fig-survey} C)).
When asked for the three most used packages, participants of the survey listed 83 packages in total.
The \emph{raster} package was mentioned the most, followed by the \emph{sf} package.
Both packages are designed for basic and advanced data handling and processing of raster and vector data, respectively, representing the results of Fig. \ref{fig:fig-survey} C).
Nevertheless, the large availability and usage of different \emph{R} packages across the community can be seen in the large ``others'' option (packages mentioned by less than 5 participants; 33.44\%).

Lastly, when asked how useful \emph{R} is currently for landscape ecology, the vast majority of participants answered with either ``very useful'' or ``useful'' (summarized 91.26\%) and only very few participants evaluated \emph{R} as ``intermediate,'' ``not useful'' or ``not useful at all'' (summarized 8.74\%; Fig. \ref{fig:fig-survey} D)).

@@

Overall, 22.3\% of the participants reported that currently no packages and functionality are missing for them or they lack the overview to answer the question.
There were three most common topics across the answers of the participants.
Firstly, many participants (13.6\%) wished for a better computational performance of \emph{R} in terms of speed and required RAM, especially for larger data.
Secondly, participants are missing specific approaches to quantify landscape characteristics (such as surface metrics), or are wishing for an improvement of currently available approaches to quantify landscape characteristics (9.7\%).
Thirdly, many participants (8.7\%) are currently missing advanced and easy-to-apply methods to create high-quality maps or other visualization-related functionality.

>>>>

To better understand how the landscape ecology community uses \emph{R}, we conducted a short survey using mailing lists and social media to reach the community.
In total, the survey was answered by 109 participants, of which the majority were either ``PhD students'' (33\%), followed by ``Post-Docs'' (26.6\%) and ``Professors'' (14.7\%).
Other, less frequent answers were ``Data scientists,'' ``None of the above,'' ``Government employees'' ``Master's degree student'' and ``Bachelor's degree student'' (in decreasing order).

Most people use \emph{R} either ``daily'' (53.2\%) or a ``few times a week'' (36.7\%).
Almost half of all participants described themselves as ``advanced'' users (46.8\%), while 40.4\% described themselves as ``intermediate'' users.
Related to this, about half of the participants either implemented their own package (21.1\%) or plan to do so in the future (22\%) and most of these packages are hosted on \emph{GitHub} and/or \emph{CRAN}.

@@

More people use the raster data model (72.5\%) in comparison to the vector data model (27.5\%).
This was also represented in the most used \emph{R} packages (Fig. \ref{fig:fig-survey} C)).
When asked for the three most used packages, participants of the survey listed 93 packages in total.
The \emph{raster} package was mentioned the most, followed by the \emph{sf} package.
Both packages are designed for basic and advanced data handling and processing of raster and vector data, respectively, representing the results of Fig. \ref{fig:fig-survey} C).
Nevertheless, the large availability and usage of different \emph{R} packages across the community can be seen in the large ``others'' option (packages mentioned by less than 5 participants; 35.87\%).

Lastly, when asked how useful \emph{R} is currently for landscape ecology, the vast majority of participants answered with either ``very useful'' or ``useful'' (summarized 90.83\%) and only very few participants evaluated \emph{R} as ``intermediate,'' ``not useful'' or ``not useful at all'' (summarized 9.17\%; Fig. \ref{fig:fig-survey} D)).

@@

Overall, 23.9\% of the participants reported that currently no packages and functionality are missing for them or they lack the overview to answer the question.
There were three most common topics across the answers of the participants.
Firstly, many participants (12.8\%) wished for a better computational performance of \emph{R} in terms of speed and required RAM, especially for larger data.
Secondly, participants are missing specific approaches to quantify landscape characteristics (such as surface metrics), or are wishing for an improvement of currently available approaches to quantify landscape characteristics (9.2\%).
Thirdly, many participants (8.3\%) are currently missing advanced and easy-to-apply methods to create high-quality maps or other visualization-related functionality.
